Does freshwater connectivity influence phosphorus retention in lakes?

### 2019-01-23 manuscript/manuscript.Rmd
formatting, concatente with appendix

liftr:
  maintainer: "Joseph Stachelek"
  email: "stachel2@msu.edu"
  from: "rocker/geospatial:latest"
  pandoc: false
  texlive: true
  sysdeps:
    - grass-dev
    - p7zip-full
    - curl
    - texlive-base
    - texlive-extra-utils 
  cran:
    - LAGOSNE
    - brms
    - dplyr
    - tidyr
    - modelr
    - data.tree
    - ggplot2
    - tidybayes
    - cowplot
    - nesRdata
    - partykit
    - ggridges
    - pinp
  remotes:
    - "jsta/nhdR"
    - "jsta/spnetwork"
    - "jsta/streamnet"
    - "jsta/rgrass7sf"

@@

# Optional: line of arbitrary text with additional information.
# Could be used, for example, to mention the bibliographic info in a post-print.
# If not specified, defaults to "This version was compiled on \today"
#date_subtitle: Published in *Journal of Statistical Software*, 2018

@@

We used data on P retention, maximum depth, and water residence time from 129 lakes in the National Eutrophication Survey \citep{usepa1975, stachelek_national_2018}. Mean annual P loading, P discharge, and P retention values in the National Eutrophication Survey (NES) dataset were calculated based on monthly sampling for P in tributary and outlet discharge points as well as any municipal waste discharges from 1972 to 1975. Here, P retention is a unitless value representing the fraction of incoming P loads. Sampling frequency for water discharge and residence time varied among lakes but details of these variations were not provided in the source dataset \citep{usepa1975, stachelek_national_2018}. Estimates of water discharge and residence time in the NES dataset represent normalized mean flow estimates expected to occur during a period of average precipitation and hydrology \citep{usepa1975}. Water residence time for our study lakes ranged from 1 week to 17 years with an interquartile range of 3 months to 1.8 years while P retention ranged from 0.06 to 0.99 with an interquartile range of 0.24 to 0.59 (Table 1).

\begin{figure} \ContinuedFloat
  \begin{center}
    \includegraphics[width=0.5\textwidth,keepaspectratio]{../tables/crop/01_lake_characteristics_table} 
    \captionsetup{labelformat=empty}
    \caption{\textbf{Table 1. } Minimum, median, maximum, and interquartile range of selected characteristics of the study lakes (N = 129).}\label{table1}    
  \end{center}
\end{figure}

We supplemented the NES dataset with boundaries for lake subwatersheds, as well as estimates of stream density, upstream lake area, upstream lake connection(s), baseflow (an index of groundwater inputs), land-use cover, and other water quality measurements from the LAGOS-NE dataset \citep[Table 1; ]{sorannoLAGOSNEMultiscaledGeospatial2017}. Our study lakes encompassed a wide range of land-use cover types and nutrient levels (Table 1). Although, lake subwatersheds were variable with respect to agricultural land use cover, we did not observe a strong relationship with lake P retention (Figure A2). On average, the water quality (total phosphorus, chlorophyll concentration, and Secchi depth) of the lakes in our study are similar to other US lakes as measured by the stratified random sampling design of the National Lakes Assessment (NLA) lake population \citep{usepa_nla2012}. However, our lakes are substantially larger and deeper than most NLA lakes (Figure A3).

@@

We modelled lake total phosphorus retention (hereafter, P retention) using the Vollenweider equation that models P retention as a function of water residence time and a parameter (k) conceptually representing in-lake P decay \citep{chapra2008surface, vollenweiderInputoutputModels1975}. Although there are several variants of this basic equation, we selected a 2-parameter form (equation \ref{eq1}) that has been
shown to have good performance in multiple cross-sectional studies \citep{brett2008, cheng2010}:

@@

where $R_i$ is P retention as a fraction of P inputs, $i$ $\tau$ is water residence time, $k$ is a unitless parameter representing in-lake P decay, and $x$ is a unitless parameter representing P export via hydrologic flushing. Here, higher values of $k$ are associated with greater integrated P losses from sedimentation and biological uptake resulting in greater P retention. Note that Eq \ref{eq1} does not include a recycling term. Therefore, our results represent net P retention (as opposed to gross P retention) under a steady state assumption where lakes are at equilibrium with respect to recycling \citep{vollenweiderInputoutputModels1975}. Note that although some forms of the Vollenweider equation use P loading as a predictor variable, it does not appear in Eq \ref{eq1}. The reason for this is two-fold. First, estimates of P loading are more difficult to obtain than estimates of water residence time and our aim was to develop a model than can be widely applied to lakes for which we lack detailed data on P loading. Second, loading based model forms have been shown to be mathematically equivalent to water residence time based model forms \citep{brett2008}.

@@

where greater differences in $k$ between R the two groups indicate greater support for a connectivity effect on P retention. Prior to model fitting, we examined the bivariate relationships between each connectivity metric, water residence time, P loading, and other factors related to P retention using Pearson’s correlation coefficients. The purpose of this exercise was twofold, to determine the potential for collinearity among any of the variables in Eq 2 and to identify any relationships between P retention, water residence time, and other watershed and lake characteristics that were not included in our model. As only one connectivity metric was used to define $g$ for each model we did not use the results of this exercise to exclude variables from further investigation. We quantified the relative support for an effect of each connectivity metric on P retention in more detail by calculating the difference in the median value of the P decay parameter $k$ between groups (i.e. $\Delta$k). We used these median $k$ values along with median estimates of $x$ to determine how differences in $k$ translated to differences in P retention (Eq 2). We judged significance by whether or not differences in group-wise P retention were greater than the measurement precision of P retention (> 0.01).

We fit all models in a Bayesian framework using the non-linear extension to the `brms` package to interface with the Stan statistical program \citep{burkner2017brms, stan2017}. In both models, we set a semi-informative prior on $k$ and $x$ of N(1.3, 0.1) and N(0.45, 0.1) respectively. These priors were based on the confidence intervals presented in \citet{brett2008} and qualitatively matched those used by \citet{cheng2010}. We used the default settings of brms and rstan to generate posterior estimates using four chains of 4,000 iterations each with no thinning and initial parameter values drawn from a uniform distribution bounded between -2 and 2. We also used the brms package for model evaluation by computing a Bayesian $R^{2}$ following the method of \citet{gelmanRsquaredBayesianRegression2017}.

@@

Our lake connectivity classes were formed by dividing the lake dataset into two classes for each connectivity metric based on the bivariate relationship between each metric and P retention. Prior studies have used a similar binary splitting approach to examine the effect of various exogenous factors on lake P retention \citep{cheng2010, shimodaIntegratingHierarchicalBayes2015}. We determined splitting criteria for each metric from the results of a random forest procedure incorporating conditional inference trees \citep[][ctree, ]{hothorn2015partykit}. This procedure creates binary splits of the independent variables (i.e. each of the connectivity metrics), which are recursively repeated to find the split that maximizes association with the dependent variable (P retention). The advantage of the ctree technique over the more typical classification-regression tree (CART) technique is that tree growth stopping rules are pre-specified \citep{hothorn2006unbiased}. As a result, some of subjectivity associated with post-hoc tree pruning is avoided.

@@

We examined the bivariate relationships between connectivity, water residence time, P loading and other factors related to P retention to determine the potential for strong relationships among any of the variables which were not accounted for by our model structure (Eq 2, Figure \ref{fig4}). We found evidence for some relationships among these variables, but none that suggest either redundancy among connectivity metrics or that otherwise limit our ability to infer relationships with P retention (Figure \ref{fig4}). For example, the Pearson correlation (r) between water residence time and stream density, which is implicitly accounted for in our model structure, was 0.29 (p < 0.05). In contrast, the correlation between water residence time and upstream lake area, which is unaccounted for in our model structure, was only 0.10 (p > 0.05). Note that interactions between connectivity metrics and water residence time are accounted for in our hierarchical model structure because Eq. 2 uses a global coefficient x for water residence time that is estimated separately from the hierarchical and connectivity-dependent P decay coefficient k. This is conceptually similar to fitting the k-connectivity-P retention relationship to the residuals of the water residence time to P retention relationship. Perhaps surprisingly, we did not observe strong correlations between P loading and water residence time (r = 0.13, p > 0.05) or between lake depth and most connectivity metrics (r < 0.17). This suggests that our P retention results are not confounded by variations in lake depth, by interactions between connectivity and P loading, or by interactions between water residence time and P loading.

@@

Differences in P retention among lakes with different connectivity metric values was reflected in differences among connectivity class-specific values of the P decay parameter $k$ (Figure 7). The connectivity metric that had the most effect on $k$ was average link length (Table 2). For instance, hierarchical models fit with the average link length metric had a median effect size of 0.23 and 0.05 for $k$ and P retention respectively (Table \ref{table2}), which means that for this metric, lakes with shorter average link lengths retained 4.7 to 4.9\% more P than lakes with longer average link lengths. The influence of lake-stream-based connectivity metrics on lake P retention was similar to stream-based connectivity metrics (Figure 7). This suggests that both lake-stream-based connectivity between lakes and streams as well as stream-based connectivity of lakes and their terrestrial watersheds influence P retention (Figure \ref{fig2}).

\begin{figure*} \ContinuedFloat
  \begin{center}
    \includegraphics[width=0.85\textwidth,keepaspectratio]{../tables/crop/02_model_results_table} 
    \captionsetup{labelformat=empty}
    \caption{\textbf{Table 2. } Connectivity class split values and samples sizes for connectivity metrics and lake depth ranked according to the difference in median $k$ (P decay parameter, $\Delta$k) values. Differences in $\Delta$k that translate to differences in P retention greater than measurement precision are marked with an asterisk. Here, WS is the lake watershed extent whereas SWS is the lake subwatershed extent. Lakes with metric values above or equal to the split value were assigned to a separate connectivity class relative to lakes below the split value. N is sample size. Connectivity metrics are defined in Figure \ref{fig2}.}\customlabel{table2}{2}    
  \end{center}
\end{figure*}

We found that connectivity metrics measured at the lake watershed extent were more strongly associated with P retention than metrics measured at the lake subwatershed extent (Table 2, Figure \ref{fig7}). Specifically, the metrics that had the strongest association with P retention such as average link length ($\Delta$k = 0.23), closest lake distance ($\Delta$k = 0.22), and stream density ($\Delta$k = 0.20) also had a stronger association at the lake watershed extent rather than at the lake subwatershed extent (Table 2, Figure \ref{fig7}). An exception to this pattern was observed with the baseflow connectivity metric where although greater differences at the lake subwatershed extent were more strongly associated with P retention, the sign of the effect was variable depending on measurement extent (i.e. the value of the P decay parameter was positively related to connectivity metric values at the subwatershed extent but was negatively related at the watershed extent). For several connectivity metrics, we judged that differences in P decay among lakes with either low or high connectivity metric values were not significant because they translated to differences in P retention that were less than measurement precision (Table 2).

@@

We found that differences in P retention among lake connectivity classes was influenced by specific connectivity metrics including average link length, closest lake distance, and stream density (Table 2, Figure \ref{fig7}). Indeed, these metrics were more strongly associated with P retention than covariates typically used in statistical P retention models (e.g. lake depth, Figure \ref{fig1}). Note that we were able to examine the influence of specific metrics as separate effects apart from water residence time because our model structure treats them as hierarchical coefficients on the P decay parameter $k$. As a result, although water residence time remains the dominant effect on lake P retention, we were able to estimate the specific influence of each metric in a way that is not possible using integrated water residence time and connectivity metrics such as watershed transport capacity \citep{fraterrigoInfluenceLandUse2008}.

@@

One of the challenges in quantifying the effect of freshwater connectivity on lake P retention is that it is not always known what spatial extent of the watershed is functionally connected to a lake. We found that connectivity at the broader lake watershed extent rather than connectivity at the finer lake subwatershed extent was more strongly associated with differences in lake P retention. In addition, we found that individual characteristics such as lake depth (or more discrete measures of connectivity such as the presence of an upstream lake) were more strongly associated with P retention than any of the other connectivity metrics measured at the lake subwatershed extent (Table 2). These findings contrast with \citet{soranno2015} who examined the association between lake nutrient concentrations (as opposed to retention) and land-use measured at varying spatial extents and found that measurements at the finer lake subwatershed extent rather than measurements at the broader lake watershed extent were more strongly associated with lake nutrient concentrations. One explanation for the difference between our results and those of \citet{soranno2015} is that connectivity metrics may reflect long-range watershed processes to a greater degree than land-use cover. An alternative explanation is that controls of lake P retention may differ compared to controls on lake P concentration.

@@

Despite our inability to test such differential transport processes across many lake watersheds at broad scales, we note that such processes are indirectly supported by our finding that connectivity metrics are associated with lake P retention. For example, differential transport of particulate matter whereby barriers to flux and differences in drainage path configuration only become apparent beyond fine spatial extents may explain why we observed stronger association of P retention with metrics measured at the broader lake watershed extent rather than metrics measured at the finer scale lake subwatershed extent. Another finding consistent with differential transport of particulate P is our observation that average link length, which approximates stream network structure and along-stream transport potential \citep{labarberaFractalDimensionStream1989}, was one of the more strongly associated metrics with lake P retention. Finally, it is notable that our stream density metric was influential at thee lake watershed extent but not at the lake subwatershed extent. Given that the stream density metric captures the average distance or drainage potential between any streams in the network and their adjacent hillslopes, floodplains, and wetlands \citep[see]{leibowitzConnectivityStreamsWetlands2018}, this suggests that differences in terrestrial runoff of particulate matter from hillslope and riparian areas are likely to be important for P retention.

>>>>

@@

# Optional: line of arbitrary text with additional information.
# Could be used, for example, to mention the bibliographic info in a post-print.
# If not specified, defaults to "This version was compiled on \today"
date_subtitle: > 
  *In Press* at *Limnology and Oceanography*, \href{https://doi.org/10.1002/lno.11137}{DOI: 10.1002/lno.11137}

@@

We used data on P retention, maximum depth, and water residence time from 129 lakes in the National Eutrophication Survey \citep{usepa1975, stachelek_national_2018}. Mean annual P loading, P discharge, and P retention values in the National Eutrophication Survey (NES) dataset were calculated based on monthly sampling for P in tributary and outlet discharge points as well as any municipal waste discharges from 1972 to 1975. Here, P retention is a unitless value representing the fraction of incoming P loads. Sampling frequency for water discharge and residence time varied among lakes but details of these variations were not provided in the source dataset \citep{usepa1975, stachelek_national_2018}. Estimates of water discharge and residence time in the NES dataset represent normalized mean flow estimates expected to occur during a period of average precipitation and hydrology \citep{usepa1975}. Water residence time for our study lakes ranged from 1 week to 17 years with an interquartile range of 3 months to 1.8 years while P retention ranged from 0.06 to 0.99 with an interquartile range of 0.24 to 0.59 (Table \ref{table1}).

\begin{figure} \ContinuedFloat
  \begin{center}
    \includegraphics[width=0.5\textwidth,keepaspectratio]{../tables/crop/01_lake_characteristics_table} 
    \captionsetup{labelformat=empty}
    \caption{\textbf{Table 1. } Minimum, median, maximum, and interquartile range of selected characteristics of the study lakes (N = 129).}\customlabel{table1}{1}
  \end{center}
\end{figure}

We supplemented the NES dataset with boundaries for lake subwatersheds, as well as estimates of stream density, upstream lake area, upstream lake connection(s), baseflow (an index of groundwater inputs), land-use cover, and other water quality measurements from the LAGOS-NE dataset \citep[Table \ref{table1}; ][]{sorannoLAGOSNEMultiscaledGeospatial2017}. Our study lakes encompassed a wide range of land-use cover types and nutrient levels (Table \ref{table1}). Although, lake subwatersheds were variable with respect to agricultural land use cover, we did not observe a strong relationship with lake P retention (Figure A2). On average, the water quality (total phosphorus, chlorophyll concentration, and Secchi depth) of the lakes in our study are similar to other US lakes as measured by the stratified random sampling design of the National Lakes Assessment (NLA) lake population \citep{usepa_nla2012}. However, our lakes are substantially larger and deeper than most NLA lakes (Figure A3).

@@

We modelled lake total phosphorus retention (hereafter, P retention) using the Vollenweider equation that models P retention as a function of water residence time and a parameter ($k$) conceptually representing in-lake P decay \citep{chapra2008surface, vollenweiderInputoutputModels1975}. Although there are several variants of this basic equation, we selected a 2-parameter form (equation \ref{eq1}) that has been
shown to have good performance in multiple cross-sectional studies \citep{brett2008, cheng2010}:

@@

where $R_i$ is P retention as a fraction of P inputs, $\tau$ is water residence time, $k$ is a unitless parameter representing in-lake P decay, and $x$ is a unitless parameter representing P export via hydrologic flushing. Here, higher values of $k$ are associated with greater integrated P losses from sedimentation and biological uptake resulting in greater P retention. Note that Eq \ref{eq1} does not include a recycling term. Therefore, our results represent net P retention (as opposed to gross P retention) under a steady state assumption where lakes are at equilibrium with respect to recycling \citep{vollenweiderInputoutputModels1975}. Note that although some forms of the Vollenweider equation use P loading as a predictor variable, it does not appear in Eq \ref{eq1}. The reason for this is two-fold. First, estimates of P loading are more difficult to obtain than estimates of water residence time and our aim was to develop a model than can be widely applied to lakes for which we lack detailed data on P loading. Second, loading based model forms have been shown to be mathematically equivalent to water residence time based model forms \citep{brett2008}.

@@

where greater differences in $k$ between the two groups indicate greater support for a connectivity effect on P retention. Prior to model fitting, we examined the bivariate relationships between each connectivity metric, water residence time, P loading, and other factors related to P retention using Pearson’s correlation coefficients. The purpose of this exercise was twofold, to determine the potential for collinearity among any of the variables in Eq 2 and to identify any relationships between P retention, water residence time, and other watershed and lake characteristics that were not included in our model. As only one connectivity metric was used to define $g$ for each model we did not use the results of this exercise to exclude variables from further investigation. We quantified the relative support for an effect of each connectivity metric on P retention in more detail by calculating the difference in the median value of the P decay parameter $k$ between groups (i.e. $\Delta$k). We used these median $k$ values along with median estimates of $x$ to determine how differences in $k$ translated to differences in P retention (Eq 2). We judged significance by whether or not differences in group-wise P retention were greater than the measurement precision of P retention (> 0.01).

We fit all models in a Bayesian framework using the non-linear extension to the `brms` package to interface with the Stan statistical program \citep{burkner2017brms, stan2017}. In both models, we set a semi-informative prior on $k$ and $x$ of N(1.3, 0.1) and N(0.45, 0.1) respectively. These priors were based on the confidence intervals presented in \citet{brett2008} and qualitatively matched those used by \citet{cheng2010}. We used the default settings of `brms` and `rstan` to generate posterior estimates using four chains of 4,000 iterations each with no thinning and initial parameter values drawn from a uniform distribution bounded between -2 and 2. We also used the `brms` package for model evaluation by computing a Bayesian $R^{2}$ following the method of \citet{gelmanRsquaredBayesianRegression2017}.

@@

Our lake connectivity classes were formed by dividing the lake dataset into two classes for each connectivity metric based on the bivariate relationship between each metric and P retention. Prior studies have used a similar binary splitting approach to examine the effect of various exogenous factors on lake P retention \citep{cheng2010, shimodaIntegratingHierarchicalBayes2015}. We determined splitting criteria for each metric from the results of a random forest procedure incorporating conditional inference trees \citep{hothorn2015partykit}. This procedure creates binary splits of the independent variables (i.e. each of the connectivity metrics), which are recursively repeated to find the split that maximizes association with the dependent variable (P retention). The advantage of the ctree technique over the more typical classification-regression tree (CART) technique is that tree growth stopping rules are pre-specified \citep{hothorn2006unbiased}. As a result, some of subjectivity associated with post-hoc tree pruning is avoided.

@@

We examined the bivariate relationships between connectivity, water residence time, P loading and other factors related to P retention to determine the potential for strong relationships among any of the variables which were not accounted for by our model structure (Eq 2, Figure \ref{fig4}). We found evidence for some relationships among these variables, but none that suggest either redundancy among connectivity metrics or that otherwise limit our ability to infer relationships with P retention (Figure \ref{fig4}). For example, the Pearson correlation (r) between water residence time and stream density, which is implicitly accounted for in our model structure, was 0.29 (p < 0.05). In contrast, the correlation between water residence time and upstream lake area, which is unaccounted for in our model structure, was only 0.10 (p > 0.05). Note that interactions between connectivity metrics and water residence time are accounted for in our hierarchical model structure because Eq. 2 uses a global coefficient $x$ for water residence time that is estimated separately from the hierarchical and connectivity-dependent P decay coefficient k. This is conceptually similar to fitting the k-connectivity-P retention relationship to the residuals of the water residence time to P retention relationship. Perhaps surprisingly, we did not observe strong correlations between P loading and water residence time (r = 0.13, p > 0.05) or between lake depth and most connectivity metrics (r < 0.17). This suggests that our P retention results are not confounded by variations in lake depth, by interactions between connectivity and P loading, or by interactions between water residence time and P loading.

@@

Differences in P retention among lakes with different connectivity metric values was reflected in differences among connectivity class-specific values of the P decay parameter $k$ (Figure \ref{fig7}). The connectivity metric that had the most effect on $k$ was average link length (Table \ref{table2}). For instance, hierarchical models fit with the average link length metric had a median effect size of 0.23 and 0.05 for $k$ and P retention respectively (Table \ref{table2}), which means that for this metric, lakes with shorter average link lengths retained 4.7 to 4.9\% more P than lakes with longer average link lengths. The influence of lake-stream-based connectivity metrics on lake P retention was similar to stream-based connectivity metrics (Figure \ref{fig7}). This suggests that both lake-stream-based connectivity between lakes and streams as well as stream-based connectivity of lakes and their terrestrial watersheds influence P retention (Figure \ref{fig2}).

\begin{figure*} \ContinuedFloat
  \begin{center}
    \includegraphics[width=0.65\textwidth,keepaspectratio]{../tables/crop/02_model_results_table} 
    \captionsetup{labelformat=empty}
    \caption{\textbf{Table 2. } Connectivity class split values and samples sizes for connectivity metrics and lake depth ranked according to the difference in median $k$ (P decay parameter, $\Delta$k) values. Differences in $\Delta$k that translate to differences in P retention greater than measurement precision are marked with an asterisk. Here, WS is the lake watershed extent whereas SWS is the lake subwatershed extent. Lakes with metric values above or equal to the split value were assigned to a separate connectivity class relative to lakes below the split value. N is sample size. Connectivity metrics are defined in Figure \ref{fig2}.}\customlabel{table2}{2}    
  \end{center}
\end{figure*}

We found that connectivity metrics measured at the lake watershed extent were more strongly associated with P retention than metrics measured at the lake subwatershed extent (Table \ref{table2}, Figure \ref{fig7}). Specifically, the metrics that had the strongest association with P retention such as average link length ($\Delta$k = 0.23), closest lake distance ($\Delta$k = 0.22), and stream density ($\Delta$k = 0.20) also had a stronger association at the lake watershed extent rather than at the lake subwatershed extent (Table \ref{table2}, Figure \ref{fig7}). An exception to this pattern was observed with the baseflow connectivity metric where although greater differences at the lake subwatershed extent were more strongly associated with P retention, the sign of the effect was variable depending on measurement extent (i.e. the value of the P decay parameter was positively related to connectivity metric values at the subwatershed extent but was negatively related at the watershed extent). For several connectivity metrics, we judged that differences in P decay among lakes with either low or high connectivity metric values were not significant because they translated to differences in P retention that were less than measurement precision (Table \ref{table2}).

@@

We found that differences in P retention among lake connectivity classes was influenced by specific connectivity metrics including average link length, closest lake distance, and stream density (Table \ref{table2}, Figure \ref{fig7}). Indeed, these metrics were more strongly associated with P retention than covariates typically used in statistical P retention models (e.g. lake depth, Figure \ref{fig1}). Note that we were able to examine the influence of specific metrics as separate effects apart from water residence time because our model structure treats them as hierarchical coefficients on the P decay parameter $k$. As a result, although water residence time remains the dominant effect on lake P retention, we were able to estimate the specific influence of each metric in a way that is not possible using integrated water residence time and connectivity metrics such as watershed transport capacity \citep{fraterrigoInfluenceLandUse2008}.

@@

One of the challenges in quantifying the effect of freshwater connectivity on lake P retention is that it is not always known what spatial extent of the watershed is functionally connected to a lake. We found that connectivity at the broader lake watershed extent rather than connectivity at the finer lake subwatershed extent was more strongly associated with differences in lake P retention. In addition, we found that individual characteristics such as lake depth (or more discrete measures of connectivity such as the presence of an upstream lake) were more strongly associated with P retention than any of the other connectivity metrics measured at the lake subwatershed extent (Table \ref{table2}). These findings contrast with \citet{soranno2015} who examined the association between lake nutrient concentrations (as opposed to retention) and land-use measured at varying spatial extents and found that measurements at the finer lake subwatershed extent rather than measurements at the broader lake watershed extent were more strongly associated with lake nutrient concentrations. One explanation for the difference between our results and those of \citet{soranno2015} is that connectivity metrics may reflect long-range watershed processes to a greater degree than land-use cover. An alternative explanation is that controls of lake P retention may differ compared to controls on lake P concentration.

@@

Despite our inability to test such differential transport processes across many lake watersheds at broad scales, we note that such processes are indirectly supported by our finding that connectivity metrics are associated with lake P retention. For example, differential transport of particulate matter whereby barriers to flux and differences in drainage path configuration only become apparent beyond fine spatial extents may explain why we observed stronger association of P retention with metrics measured at the broader lake watershed extent rather than metrics measured at the finer scale lake subwatershed extent. Another finding consistent with differential transport of particulate P is our observation that average link length, which approximates stream network structure and along-stream transport potential \citep{labarberaFractalDimensionStream1989}, was one of the more strongly associated metrics with lake P retention. Finally, it is notable that our stream density metric was influential at the lake watershed extent but not at the lake subwatershed extent. Given that the stream density metric captures the average distance or drainage potential between any streams in the network and their adjacent hillslopes, floodplains, and wetlands \citep[see ][]{leibowitzConnectivityStreamsWetlands2018}, this suggests that differences in terrestrial runoff of particulate matter from hillslope and riparian areas are likely to be important for P retention.
